Data Management for LSST Special Programs

### 2023-10-18 sec_proc.tex
updated through SP

\textbf{The meaning of ``can be processed".}
The condition ``can be processed" is ultimately left to the discretion of
Rubin Data Management in Rubin Operations, but is expected to include
all standard and alternative visits in sky regions for which a template image exists.
This might also include some non-standard visits (shorter or longer exposures), 
as long as they can be processed by the Prompt pipeline and an appropriate template image exists.
Visits with very short or very long exposure times (or very bright sky 
backgrounds) might be excluded if they would need specialized algorithms for,
e.g., instrument signature removal, difference-imaging, template-generation 
(Section~\ref{ssec:proc_bounds}).

@@

\textbf{WFD and Special Program data would co-reside.}
No ``unique and separate" data products for the Special Progams would be 
produced by regular Prompt processing.
Special Programs data that is processed by the Prompt pipeline would 
contribute to the Prompt data products for the WFD program as 
described in Section 3 of the DPDD \citedsp{lse-163}. 
These data products are the results of Difference Image Analysis (DIA),
such as the difference images, catalogs of sources detected in difference
images ({\tt DiaSources}) and associated static-sky {\tt DiaSources}
into {\tt DiaObjects}, and alert packets.

@@

\item {\tt DiaObject} histories may become too large for the sizing model.
Alert packets contain the full records of all associated 
{\tt DiaSources} from the past 12 months, but the alert
stream bandwidth is sized for the expected histories for
WFD program fields. 
The Prompt pipeline resources are also sized for the
WFD program, and it might not be possible to load up
thousands of epochs at a time.
The Data Management team will have to test the realized
capabilities of Prompt processing and alert distribtuion,
and potentially impose a mitigation strategy such as
limiting histories to the last $N$ observations instead
of the last 12 months in heavily-observed regions.

\item The association of {\it new} {\tt DiaSources} into {\tt DiaObjects} 
may be compromised.
For consecutive images, processing for the second image begins when the processing for the 
first image is only halfway done.
At this point, the {\tt DiaObject} catalog has 
not yet been updated with the new {\tt DiaSources} detected in the first image.
Thus, the {\tt DiaSource}s from images one and two for a new transient 
would not be associated with a single {\tt DiaObject}, but instead would 
each instantiate a new {\tt DiaObject}.

@@

\textbf{Difference images, {\tt DiaSource} and {\tt DiaObject} catalogs -- }
As all visits can contribute to the time-domain astronomy science goals 
of the LSST, all Special Programs data that {\it can} be processed with 
the difference image analysis (DIA) pipeline alongside the WFD visits 
{\it should} be.
All of this processing should use the same template image for a given
field.
Users who want, e.g., only {\tt DiaSources} a given DDF, would be 
able to write queries using the region and program labels in the catalog.

\textbf{Processed visit images, {\tt Source} and {\tt ForcedSource} catalogs -- }
For the same reasons as above, all Special Programs data that 
{\it can} be processed with the Data Release pipieline alongside the WFD 
visits {\it should} be.

\textbf{Deep coadds and the {\tt Object} catalog -- }
Some of the core LSST science goals require a WFD main survey data 
products of nearly uniform depth.
Whether and how to include any Special Programs data in the deep image 
coadds and the corresponding {\tt Object} catalog is left entirely to the 
discretion of the RDM team in Operations.

For example, perhaps Special Programs images will only be included when 
they bring additional area to the same depth as the rest of the WFD, or 
when they suppress edge effects or low-order modes in the 
all-sky photometric solutions.

For stacked images and their associated catalogs, the region labels should be
included (but the program labels would not be needed or possible).

@@

In short, \emph{possible} means that original or reconfigured versions of the LSST
Science Pipelines can be used , and \emph{necessary} means the primary science goal for Special Program
could not be met without the data products produced by the Special Processing.
The definitions of possible and necessary, and the scope of Special Processing,
are ultimately left to the discretion of Rubin Data Management in Operations.

@@

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as ``not necessary" and requiring 
User-Generated Processing.

@@

\textbf{Tables should be joinable -- } where feasible.

\textbf{A note about intermediate-depth ``custom" coadds -- }
As with the annual WFD deeply coadded images and associated catalogs, RDM would 
only generate one coadd (including all images to date) for a given special region.
Custom coadds (e.g., weekly, monthly) for very specific science goals are considered 
user-generated data products, for both WFD and Special Programs.

As the DDFs will likely be observed with standard or alternative standard 
visits, it is likely that RDM will be able to reconfigure existing pipelines 
to process the DDF data and produce ``unique and separate" DDF data products.

For example, RDM-generated ``unique and separate" DDF data products might 
include:
\begin{itemize}
\item nightly-coadded images (24 h)
\item nightly-coadded difference images (24 h)
\item {\tt DiaSource}- and {\tt DiaObject}-like catalogs for the nightly-coadds (24 h)
\item deeply-coadded images (all images to date; yearly)
\item templates for the nightly-coadded difference images (yearly)
\item {\tt Source}- and {\tt Object}-like catalogs for the nightly-coadded and deeply-coadded images (yearly)
\end{itemize}

@@

As of the Phase 2 SCOC recommendations \citedsp{PSTN-055}, three kinds of 
short-exposure mini-surveys were under consideration: 
a static short exposure map of the sky in $ugrizy$ for calibration; 
a static to transient short exposure survey; and 
an Near-Earth Objects (NEO) twilight survey.

If either the transients or NEO short-exposure suvey is executed, and if 
RDM finds that the LSST Science Pipelines and, specifically, template images 
can be made and difference image analysis (DIA) run, then the RDM-generated 
data products may be similar to the Prompt data products (potentially 
including alerts; Section 3 of \citeds{lse-163}).

If either of the static short-exposure mini-surveys is executed, and if 
RDM finds that the LSST Science Pipelines can be reconfigured and used 
for short exposures, then the RDM-generated data products would likely be 
of a similar format to the Data Release data products 
(Section 4, \citeds{lse-163}).
For example, a separate repository of processed visit images and coadded 
images, and separate (but joinable) {\tt Source}, {\tt ForcedSource}, and 
{\tt Object} catalogs.

The cases where ``unique and separate" data products for mini-surveys are 
needed to reach the science goals of a Special Program, and are possible for 
RDM to generate, are most likely to be of a similar format to the annual 
Data Release data products: static-sky coadded images and associated 
{\tt Source}- and {\tt Object}-like cataogs for standard or alternative visits.

In cases where the observations are all obtained with a short season 
(e.g., a week spent observing a specific region), such ``annual" 
data products might be generated and released on an intermediate time scale, 
if resources allow.

>>>>

\textbf{The meaning of ``can be processed".}
This is ultimately left to the discretion of
Rubin Data Management in Rubin Operations, but is expected to include
all standard and alternative visits in sky regions for which a template image exists.
This might also include some non-standard visits (shorter or longer exposures), 
as long as they can be processed by the Prompt pipeline and an appropriate template image exists.
Visits with very short or very long exposure times (or very bright sky 
backgrounds) might be excluded if they would need specialized algorithms for,
e.g., instrument signature removal, difference-imaging, template-generation 
(Section~\ref{ssec:proc_bounds}).

@@

\textbf{How WFD and Special Program data would co-reside.}
No ``unique and separate" data products for the Special Progams would be 
produced by regular Prompt processing.
Special Programs data that is processed by the Prompt pipeline would 
contribute to the Prompt data products for the WFD program as 
described in Section 3 of the DPDD \citedsp{lse-163}. 
These data products are the results of Difference Image Analysis (DIA),
such as the difference images, catalogs of sources detected in difference
images ({\tt DiaSources}) and associated static-sky {\tt DiaSources}
into {\tt DiaObjects}, and alert packets.

@@

\item \textbf{{\tt DiaObject} histories may become too large for the sizing model.}
Alert packets contain the full records of all associated 
{\tt DiaSources} from the past 12 months, but the alert
stream bandwidth is sized for the expected histories for
WFD program fields. 
The Prompt pipeline resources are also sized for the
WFD program, and it might not be possible to load up
thousands of epochs at a time.
The Data Management team will have to test the realized
capabilities of Prompt processing and alert distribtuion,
and potentially impose a mitigation strategy such as
limiting histories to the last $N$ observations instead
of the last 12 months in heavily-observed regions.

\item \textbf{The association of {\it new} {\tt DiaSources} into {\tt DiaObjects} 
may be compromised.}
For consecutive images, processing for the second image begins when the processing for the 
first image is only halfway done.
At this point, the {\tt DiaObject} catalog has 
not yet been updated with the new {\tt DiaSources} detected in the first image.
Thus, the {\tt DiaSource}s from images one and two for a new transient 
would not be associated with a single {\tt DiaObject}, but instead would 
each instantiate a new {\tt DiaObject}.

@@

\textbf{Time-domain DIA data products --}
This includes the results of the annual reprocessing of WFD program data with 
Difference Image Analysis (DIA), and the production of Data Release
versions of the processed images (single-visit and difference images)
and associated catalogs ({\tt DiaSource}, {\tt DiaObject}, {\tt Source},
{\tt ForcedSource}, {\tt DiaForcedSource}, and so on).
These data products will be used primarily for time-domain science.
They should include Special Programs data for the same reasons as
provided in Section~\ref{ssec:sci_pproc}, and with the same 
considerations as discussed in Section~\ref{sssec:proc_reg_prompt}.
This processing should use the same template image for a given field.

\textbf{Static-sky data products --}
This includes the tessellation and coaddition of WFD program images
and the associated multi-band {\tt Object} catalog and survey property maps.
Whether and how to include any Special Programs data in these data products
is left entirely to the discretion of the Rubin Data Management team in Operations.
As an example, perhaps Special Programs images will only be included when they 
assist with uniformity or suppress edge effects or low-order modes in the 
all-sky photometric solutions.

@@

In short, \emph{possible} means that original or reconfigured versions of the LSST
Science Pipelines can be used , and \emph{necessary} means the primary science goal for Special Program
could not be met without the data products produced by the Special Processing.

\textbf{Joinable tables --}
Any tables for the unique and separate data products should be joinable to the 
data products for the WFD program, when possible.

\textbf{Survey property maps --}
As for the tesslated all-sky coadded images made from WFD program data, survey property maps
should be made individually as part of any Special Programs that generates
tesselated coadded images.

\textbf{Special Processing timescales --}
The timescales for Special Processing should be adopted that best serve the
primary science goal for the Special Program.
For example, nighly-coadd difference-image analysis for DDFs is the most
scientifically useful if done on a daily cadence, but the deeply coadded
images for DDF fields could be released annually.

The interpretations of possible and necessary, and the scope of Special Processing,
are ultimately left to the discretion of Rubin Data Management in Operations.

@@

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as ``not necessary" and requiring 
User-Generated Processing.
Note that for the WFD program data, such custom coadds are also considered 
as beyond scope and in need of User-Generated Processing.

@@

Further examples of Special Processing are provided below.

\subsubsection{Deep Drilling Fields (DDFs)}

As the DDFs will likely be observed with standard or alternative standard 
visits, Data Management will be able to reconfigure existing pipelines for
Special Processing to produce unique and separate DDF data products.

For example, Special Processing for the DDF data products might include:
\begin{itemize}
\item nightly-coadded images (24 h)
\item nightly-coadded difference images (24 h)
\item {\tt DiaSource}- and {\tt DiaObject}-like catalogs for the nightly-coadds (24 h)
\item deeply-coadded images (all images to date; yearly)
\item templates for the nightly-coadded difference images (yearly)
\item {\tt Source}- and {\tt Object}-like catalogs for the nightly-coadded and deeply-coadded images (yearly)
\end{itemize}

@@

As described in \citedsp{PSTN-055}, there are a few
short-exposure mini-surveys are under consideration.
Two examples are a short exposure map of the sky in $ugrizy$ for calibration,
and a Near-Earth Objects (NEO) twilight survey.

Special Processing for short-exposure mini-surveys remains to
be determined.
Since it falls under the remit of Data Management to perform proper calibration,
an evaluatation of whether short exposures for calibration are necessary
will be done. 
If Data Management does find that short-exposure and/or high-sky brightness images
can be processed with reconfigured versions of the LSST Science Pipelines,
then unique and separate data products could be generated with Special Processing.
For the NEO twilight survey, these data products would likely be similar to the
Prompt or DIA data products.
For the calibration survey, these data products would like be similar to the
annual data release tessellated coadds and associated catalog.

\subsubsection{Standard Visit Mini-Surveys}

Consider a Special Program in which a special region of sky is observed with standard
visits but a a special strategy or cadence which is significantly distinct from the WFD program,
and lasts for a limited amount of time.
For example, a short-term survey of the Magellanic Clouds, with dual primary science
goals in time-domain and static-sky science.

In cases like this, a set of unique and separate data products with the same formats as the 
time-domain DIA and static-sky data products described in Section~\ref{sssec:proc_reg_dr}
should be created with Special Processing.
They might be released with an annual data release or on an intermediate timescale, e.g.,
within six months of the conclusion of the mini-survey observations.
